### 2022-12-23 paper/boneyard.tex
added some potentially useful text from a former grant application

% text copied from early NSF grant draft (unused in final grant)

Memory researchers, like cognitive neuropsychologists in many domains, have historically taken a ``less is more'' approach.  They follow the logic that, given a memory topic of interest (e.g.\ memory for real-world autobiographical events), we can ask: what is the simplest experimental paradigm that we might reasonably expect to involve a shared or analogous set of behavioral tendencies and brain structures?  There are many benefits to taking a simplified approach to studying complex cognitive phenomena.  By composing a set of experimental circumstances where every aspect of the stimuli and timing is carefully controlled by the experimentalist, and where the behaviors a participant can exhibit are highly constrained, it becomes possible to systematically link specific experimental observations to their corresponding neural and behavioral mechanisms~\citep{OReiMuna00}.  However, there is also a major cost to simplifying: tightly controlled experiences designed in the lab can often only provide indirect insights into the real-world analogs of those experiences.  This has led to a major hole in the memory theory literature.  On the one hand, the most well-developed theories of human memory have been informed by a long history of tightly controlled (but simplified relative to the real-world) experiments~\citep[e.g.\ list-learning, priming, reaction times; ][]{Kaha12}.  On the other hand, studies of real-world classroom learning have largely avoided drawing direct links with those theories, instead focusing on questions about applied topics such as creating effective assessment tools, increasing student engagement and motivation, and understanding what makes an effective teacher~\citep{Seid07}.

The general approach we take in our project is to use prior experimental and theoretical work on list learning as a foundation for optimizing STEM learning.  Specifically, \textbf{Module 1 is centered around developing approaches for improving participants' memories for lists of words-- a simple and well-studied task expected to involve neural mechanisms similar to those underlying STEM learning}.  To foreshadow our proposal, \textbf{in Module 2, we have designed a set of analogous experiments and analyses that adapt the approach in Module 1 to a more naturalistic learning setting (learning from online STEM training videos)}.  By mirroring our approaches in Modules 1 and 2, we hope to be able to draw parallels between the two sets of experiments and findings.  Finally, \textbf{in Module 3, we plan to replicate the EEG experiments in Modules 1 and 2 using a low-cost EEG system that our team developed.  We will also develop (and document) an open-source toolkit for sharing our software (and data).  In this way, Module 3 will serve to help bring our work into real-world classrooms, and to real-world students}.

\subsection*{Context and the structure of memory}
Remembering words on a studied list requires distinguishing the current list from the rest of one's experience.  To model this fundamental memory capability, cognitive scientists have posited the existence of a special representation, called \emph{context}, that is associated with each list.  According to early theories~\citep[e.g.][]{Este55a,AndeBowe72} context representations are composed of many features which fluctuate from moment to moment, slowly drifting through a multidimensional feature space.  During recall, this representation forms part of the retrieval cue, enabling us to distinguish list items from non-list items.  Understanding the role of context in memory processes is particularly important in self-cued memory tasks, such as \textit{free recall}, where the retrieval cue is ``context'' itself.  In free recall, participants study lists of items and are instructed to recall the items in any order they choose.

\begin{wrapfigure}[10]{r}{1.3in}
  \begin{center}
    \vspace{-33pt}
    \includegraphics[width=1.1in]{figs/fingerprint_schematic.pdf}
  \end{center}
  \vspace{-18pt}
  \caption{\footnotesize \textbf{Memory fingerprint.}  Reflects a participant's tendency to cluster recalls according to each stimulus feature dimension.}
  \label{fig:fingerprint}
\end{wrapfigure}

Over the past half-century, context-based models have enjoyed impressive success at explaining many stereotyped behaviors observed during free recall and other list-learning tasks \citep[e.g.][]{Este55a, RaaiShif80, GlenEtal83, HowaKaha02a, SiroEtal05, KimbEtal07, PolyKaha08, PolyEtalTulv, SedeEtal08, PolyEtal09, ShanHowa12}.  These phenomena include the well-known recency and primacy effects (superior recall of items from the end and, to a lesser extent, from the beginning of the study list), as well as semantic and temporal clustering effects \citep[for review see][]{KahaEtal08}.  The contiguity effect is an example of temporal clustering, which is perhaps the dominant form of organization in free recall.  This effect can be seen in the tendency for people to successively recall items that occupied neighboring positions in the study list.  For example, if a list contained the sub-sequence \textsc{``absence hollow pupil''} and the participant recalls the word \textsc{``hollow''}, it is far more likely that the next response will be either \textsc{``pupil''} or \textsc{``absence''} than some other list item~\citep{Kaha96}.  In addition, there is a strong forward bias in the contiguity effect: subjects make forward transitions (i.e., \textsc{``hollow''} followed by \textsc{``pupil''}) about twice as often as they make backward transitions, despite an overall tendency to begin recall at the end of the list.  There are also striking effects of semantic clustering \citep{RomnEtal93, Bous53, BousEtal54, JenkRuss52, MannKaha12}, whereby the recall of a given item is more likely to be followed by recall of a similar or related item than a dissimilar or unrelated one.

In general, people organize memories for words along a wide variety of stimulus dimensions.  As captured by models like the \textit{Context Maintenance and Retrieval Model}~\citep{PolyEtal09}, the stimulus features associated with each word (e.g.\ the word's meaning, font size, font color, location on the screen, size of the object the word represents, etc.) are incorporated into the participant's mental context representation~\citep[also see][]{SmitVela01, MannEtal15}.  During a memory test, any of these features may serve as a memory cue, which in turn leads the participant to recall in succession words that share stimulus features.  We refer to a participant's particular tendency to organize their recalls according to each stimulus dimension as their \textit{memory fingerprint} (Fig.~\ref{fig:fingerprint}).  Formally, the memory fingerprint is a vector whose elements range between 0 and 1, where each element of the memory fingerprint reflects a stimulus feature.  A value close to 1 indicates that the participant nearly always orders their recalls according to the given stimulus dimension.  A value close to 0 indicates that the participant almost \textit{never} orders their recalls according to the given stimulus dimension; a value close to 0.5 indicates that the participant's recalls are random with respect to that stimulus dimension.  In Experiment 1 we propose a free recall variant where each stimulus has many well-defined and quantifiable features, enabling us to track how the ways people organize their memories (i.e., their memory fingerprints) change over time.  In Experiment 2 we use a behavior-driven approach to manipulate these processes, and in Experiment 3 we use electroencephalographic (EEG) data to manipulate these processes.  In this way, Experiment 1 is focused on elucidating the dynamics of this very specific form of learning, and Experiments 2 and 3 are focused on manipulating the learning process, with the goal of improving (or diminishing) learning efficiency.

\subsection*{The feature-rich free recall paradigm}

\begin{wrapfigure}[10]{r}{1.4in}
\vspace{-57pt}
  \begin{center}
    \includegraphics[width=1.2in]{figs/FRFR}
  \end{center}
\vspace{-16pt}
  \caption{\footnotesize \textbf{Feature-rich free recall.}  After studying lists comprised of words that vary along several feature dimensions, participants verbally recall words in any order (microphone icon).}
  \label{fig:exp1}
\end{wrapfigure}

Each of our Module 1 experiments is based on a variant of the classic free recall paradigm that we term \textit{feature-rich free recall}.   In feature-rich free recall, participants study 16 lists, each comprised of 16 words that vary along a number of stimulus dimensions: semantic category, object size, text color, text location, word length, and starting letter (Fig.~\ref{fig:exp1}). Each list contains four words from each of four different semantic categories; all other stimulus features are randomized.  Each word is presented for 2 seconds, followed by a 2 second pause (blank screen).  After studying each list, the participant has 60 seconds to recall as many words as they can from that list, in any order they choose.  Because each individual word is associated with several well-defined (and quantifiable) features, and because each list incorporates a diverse mix of feature values along each dimension, this allows us to evaluate participants' memory fingerprints in rich detail.

Our experimental paradigm incorporates the Google Cloud Speech API speech-to-text engine to automatically transcribe participants' verbal recalls into text.  This allows recalls to be transcribed in real time-- a distinguishing feature of the experiment that is critical for using participants' recalls of prior lists to adapt the presentation orders of future lists, as in Experiments 2 and 3.  (In typical verbal recall experiments the audio data must be parsed manually, precluding this sort of design.)  We recently published a study verifying that the computer-generated recall transcripts are sufficiently similar to human-generated transcripts to yield reliable data~\citep{ZimaEtal18}.

\subsection*{Experiment 1: tracking the dynamics of memory organization over successive learning trials}
% maybe this could benefit from reframing wrt to the variance idea?
Large portions of the memory theory and education literatures have focused on identifying learners' unique learning styles.  In the memory theory literature~\citep{SmitVela01, Kaha12, HealKaha14a, MannEtal15}, the focus has been on identifying behavioral signatures of how memories are organized (i.e., in the language of this proposal, identifying memory fingerprints).  In the education literature, the focus has been been on asking whether catering lessons to an individual's unique learning style can help them to learn more efficiently~\citep{PashEtal08}.  Each of these literatures effectively treats people's learning styles as a fixed point.  \textbf{In other words, the assumption underlying the majority of these studies is that each individual has their own unique learning style, and that learning style remains roughly consistent over time.  Experiment 1 is designed to explicitly test this assumption using the feature-rich free recall paradigm (Fig.~\ref{fig:exp1}).}  Importantly, we will collect data from two populations: (a) members of the Dartmouth and surrounding community, with the experiment running in a tightly controlled laboratory setting, and (b) Mechanical Turk workers, via the Internet.  Taken together, we hope that these two populations of participants will allow us to generalize our findings to a diverse set of learners.

\paragraph{Design and procedure.}
We will use the feature-rich free recall paradigm (Fig.~\ref{fig:exp1}) to study how people's memory fingerprints are influenced by presentation order.  The experiment will be divided into six manipulation conditions (one for each stimulus feature); in each of these manipulation conditions the words from the first 8 lists participants study will be sorted according to one of the stimulus dimensions.  For example, in the ``color'' condition, words from the first 8 lists will be sorted by font color; in the ``starting letter'' condition, words from the first 8 lists will be sorted alphabetically; etc.  The second 8 lists will be presented in a random order in every condition.  As described below, we will assess how participants' memories are affected by these order manipulations, both for their memories of words on first 8 lists where the manipulations are applied, as well as on the second 8 lists.  Participants' behaviors will be compared to a seventh (control) condition, where the presentation order is randomized for all 16 lists.

We carried out a power analysis (power: 0.8, Type 1 error rate: 0.05) using our prior work to estimate the relevant parameters~\citep{MannEtal11, MannEtal12}.  The analysis indicated that we will require 30 in-laboratory participants in each of the six manipulation conditions to achieve sufficient power.  Because all of our findings will be compared relative to the random control condition, we plan to run 60 in-laboratory participants in that condition.  Prior work suggests that Mechanical Turk data is more variable than data collected in the laboratory~\citep{CrumEtal13}, so in each condition we plan to collect data from twice as many Mechanical Turk workers as in-laboratory participants.  In all, across the 7 experimental conditions, we plan to run 720 participants in Experiment 1 (240 in the lab and 480 online through Mechanical Turk).

\paragraph{Key analyses.}
Experiment 1 is motivated by two fundamental questions about the dynamics of memory organization and recall.  \textbf{First, we want to understand how memory fingerprints vary across people and over time.}  For example, are people's memory fingerprints generally stable across lists?  Or, what are the similarities and differences across people (in how people organize their memories, and how their memory fingerprints change over time)?  Are those individual differences predictive of memory performance? \textbf{Second, we want to understand how people's memory fingerprints can be manipulated.}  Specifically, if lists are sorted according to a particular feature (e.g. the words' semantic categories), how does that influence people's tendencies to cluster their recalls of those words with respect to that feature?  And which of those tendencies persist during recalls of later (randomly ordered) lists?  For example, if we organize early lists alphabetically, are  participants more likely recall the words from those lists alphabetically as well?  More generally, might there be some way of influencing how people organize their memories, even after the influencing cue is no longer present?

% figures: pilot data displaying random-condition fingerprints, relationship between fingerprint variance and memory
\begin{wrapfigure}[10]{R}{4.0in}
  \vspace{-52pt}
  \begin{center}
    \includegraphics[width=4.0in]{figs/random_pilot.pdf}
  \end{center}
\vspace{-18pt}
  \caption{\footnotesize \textbf{Preliminary data for feature-rich free recall.}  \textbf{a.} Clustering scores along each feature dimensions (i.e. memory fingerprints).  Each dot represents the average clustering score along one dimension, for one participant. \textbf{b.} Across-subject correlation between fingerprint variability and memory performance.  \textbf{c.} Across-condition correlation between fingerprint variability and memory performance.}
  \label{fig:random_pilot}
\end{wrapfigure}

\paragraph{Experiment 1 preliminary data.}
We ran a preliminary study ($n = 172$) to measure participants' memory fingerprints under the in-laboratory conditions outlined in Experiment 1.  Using data from the control condition (where all lists were ordered randomly; $n = 45$), we computed the distribution of average memory fingerprints (across lists) from each participant (Fig.~\ref{fig:random_pilot}a).  The degree to which participants organize their memories with respect to each feature varies substantially across people (e.g.\ there is a wide spread of values covered across participants within each feature dimension).  In addition, some features are more likely to be leveraged in people's memories than others.  For example, participants are more likely to organize their recalls by category, size, and temporal proximity than by font color or word length.  We also examined how stable participants' memory fingerprints were across lists.  We found that participants whose memory fingerprints were more stable across lists (i.e., lower variance) remembered more words (Fig.~\ref{fig:random_pilot}b).  We also examined the manipulation conditions, where the first 8 lists participants studied were sorted according to one feature (category: $n=21$, color: $n=20$, screen location: $n=21$, first letter: $n=20$, word length: $n=23$, size: $n=22$).  Relative to randomly sorted lists, sorting by some features increased the average memory fingerprint variability and sorting by other features decreased the average memory fingerprint variability.  (We used a permutation-based procedure to compute the variability, which controlled for the distributions of feature values of the words participants remembered.)  We found that participants remembered more words in the conditions where the presentation orders led to lower fingerprint variability (Fig.~\ref{fig:random_pilot}c).  Taken together, these preliminary data suggest that participants who naturally exhibit more stable fingerprints remember more words, and fingerprint stability may be influenced by the order in which words are presented.  In turn, sorting the word presentations along different dimensions can enhance (or diminish) memory performance according to whether fingerprint variability is increased or decreased, respectively.

\subsection*{Experiment 2: adaptive list learning using (behavioral) memory fingerprints}
\begin{wrapfigure}[9]{R}{1.6in}
  \vspace{-25pt}
  \begin{center}
    \includegraphics[width=1.5in]{figs/adaptive_fr_flowchart.pdf}
  \end{center}
\vspace{-20pt}
  \caption{\footnotesize \textbf{AdaptiveFR.}  The average memory fingerprint is updated after recalling each list, and is used to determine the presentation order of the next list.}
  \label{fig:adaptiveFR_flowchart}
\end{wrapfigure}

Our pilot data from Experiment 1 suggest that the order in which words are studied affects how they are later recalled.  Further, certain orderings can lead to lower memory fingerprint variability, which in turn leads (on average) to better memory for the studied words.  Whereas Experiment 1 explores the memory effects of pre-defined order manipulations, in Experiment 2 we propose to adapt the presentation order of future lists, using behavior measured on previous lists.  \textbf{Our overarching goal is to explore whether memory models that are custom-fit to individual participants may be exploited (through presentation order manipulations) to enhance or diminish memory performance.}

\begin{wrapfigure}[24]{R}{2in}
  \vspace{-25pt}
  \begin{center}
    \includegraphics[width=2in]{figs/adaptive_pilot}
  \end{center}
\vspace{-20pt}
  \caption{\footnotesize \textbf{AdaptiveFR pilot data. a.} Fingerprint variability decreases when lists are sorted according to the average memory fingerprints from prior lists (stabilize condition).  \textbf{b.}  Stabilizing memory fingerprints improves recall specifically for middle-of-the-list words.}
  \label{fig:adaptive_pilot}
\end{wrapfigure}

\paragraph{Design and procedure.} We will use a modified version of the feature-rich free recall paradigm designed to enhance or diminish participants' memory performance by adaptively adjusting the presentation order of future lists using recall data from previous lists (\textit{AdaptiveFR}; Fig.~\ref{fig:adaptiveFR_flowchart}).  As in Experiment 1, participants in Experiment 2 will study 16 lists.  The first 4 lists will be sorted randomly; these are used to compute the participants' individualized memory fingerprints.  Each of the remaining 12 lists (grouped into three blocks of 4 lists) will be sorted according to one of three algorithms (every participant will study and recall lists sorted using all three algorithms, but in a block-randomized order).   In the \textit{stabilize} condition, the list words will be sorted into an order that matches the participant's average memory fingerprint from all prior lists.  For example, if the memory fingerprint indicates a strong tendency to cluster recalls by category, words from the same category will tend to appear at adjacent study positions on ``stabilize'' lists.  To the extent to which the presentation order influences recall order, the stabilize condition should reduce variability in participants' memory fingerprints (by influencing future memory fingerprints to match the current average fingerprint). In the \textit{destabilize} condition, words will be sorted in an attempt to maximize the fingerprint variability.  Specifically, we will generate 10,000 random permutations of each to-be-presented list, compute a memory fingerprint for each (treating each permutation as if it were a recall sequence), and select the permutation whose fingerprint is most different (in terms of Mahalanobis distance) from the average fingerprint from prior lists.  Finally, in the \textit{random} condition, words will be shuffled randomly; this condition will be used as a baseline for the others.  We also plan to run series of analogous yoked control conditions in a second set of participants, whereby memory fingerprints based on \textit{other} participants' recalls are used to sort future lists.

We carried out a power analysis (power: 0.8, Type 1 error rate: 0.05) using our pilot data as a guide (Fig.~\ref{fig:adaptive_pilot}).  The analysis indicates that we will require at least 33 in-laboratory participants to achieve sufficient power, which we plan to round up to 40 participants.  Similarly (following our logic from Experiment 1) we plan to run 80 Mechanical Turk workers in the online version of Experiment 2.  We will also duplicate the experiment using yoked controls, for a total of 80 in-laboratory and 160 Mechanical Turk participants.

\paragraph{Key analyses.}  The primary questions driving Experiment 2 concern how list order manipulations can influence memory.  \textbf{First, we want to understand whether we can adaptively enhance or diminish memory performance by increasing or decreasing the stability of people's memory fingerprints.}  Essentially, Experiment 2 is intended to serve as a strong test of our hypothesized causal link between the stability of people's memory fingerprints and their memory performance.  We will test whether our experimental manipulations can indeed influence fingerprint stability, and if so, whether those manipulations also impact memory performance.  \textbf{Second, we want to understand the extent to which these person-specific reordering manipulations are beneficial, relative to population-driven reordering manipulations.}  For example, if we customize the word presentation orders based on an individual's data, do we see stronger effects on their performance than if we had instead customized the lists based on the average data from other participants?

\paragraph{Pilot data.}  We ran a pilot study ($n = 12$) to test the feasibility of our approach.  We found that when word lists were sorted to be consistent with a participant's memory fingerprint (the stabilize condition), this led to lower fingerprint variability ($t(11) = 1.94, p = 0.06$; Fig.~\ref{fig:adaptive_pilot}a) and enhanced memory (Fig.~\ref{fig:adaptive_pilot}b).  Notably, this memory enhancement was specific to words from the middle of the list (positions 8--12; $t(11)=3.18, p=0.008$). This provides preliminary evidence that manipulations that increase the stability of participants' recall tendencies also improve memory.

\subsection*{Experiment 3: adaptive list learning using (neural) memory fingerprints}
The above experiments leverage participants' \textit{behaviors} (i.e., the sequences of recalls they make) to optimize learning by changing the order in which subsequent lists are presented.  In Experiment 3, we seek to optimize learning using \textit{neural} (EEG) data. \textbf{Our prior work shows that the neural patterns recorded while participants study a list of words may be used to predict their individual tendencies to cluster their recalls according to presentation order (temporal clustering; Fig.~\ref{fig:clust}a) or word meaning (semantic clustering; Fig.~\ref{fig:clust}b).}

% Logically, there are two sets of factors that can contribute to recall order.  The first set of factors are processes that occur at the time information is encoded.  The results summarized in Figure~\ref{fig:clust} show that some information about the order in which words on a studied list will later be recalled may be predicted from neural recordings taken at the time of the words are studied (encoded).  The second set of factors that can contribute to recall order occur at retrieval.  Any additional processes or stimuli that occur as the participant recalls the words could conceivably affect recall order.  For example, a sudden pang of hunger or boredom that happens to occur during the recall phase of the experiment might lead the participant to recall other words associated with those new thoughts.  The neural patterns recorded as the participant studies a list of words cannot be used to recover order effects driven by retrieval-specific processes.

\begin{wrapfigure}[12]{R}{2.5in}
\vspace{-32pt}
  \begin{center}
    \includegraphics[width=2.5in]{figs/neural_predictors_of_clustering.pdf}
  \end{center}
\vspace{-19pt}
  \caption{\footnotesize \textbf{Pilot data: neural signatures of memory fingerprints.}  Electrophysiological data recorded during study may be used to predict the degree to which a participant will later (\textbf{a.}) temporally cluster their recalls; panel adapted from~\citep{MannEtal11} or (\textbf{b.}) semantically clustering their recalls; panel adapted from~\citep{MannEtal12}.}
  \label{fig:clust}
\end{wrapfigure}

In Experiment 3, we ask whether these prior findings extend to additional dimensions of the memory fingerprint (i.e., how participants order their recalls of words).  Following from our prior work, we will record EEG activity (wavelet-derived mean spectral power in 50 log-spaced frequency bands, from each electrode) as each word is studied~\citep{MannEtal11, MannEtal12}.  We will also examine patterns of cross-frequency coupling and other dynamic connectivity patterns~\citep{MannEtal14b}.  We will then identify a neural predictor for each dimension of the memory fingerprint-- i.e., one neurally derived predictor each of how likely the word being studied will follow a word matching the same semantic category, or with a similar color, location, starting letter, etc.  Specifically, to obtain each predictor (e.g.\ of clustering recalls by color), we will consider the set of correctly recalled words (excluding the first).  For each recall in turn, we will ask: considering the dimension of interest (e.g.\ font color), how similar is each not-yet-recalled word to the previous recall?  We will use the distribution of similarity values to tag each correct recall with a percentile rank.  Regressing the neural activity recorded just before each word was studied against these percentile ranks will yield a neural predictor (classifier) of clustering by the given memory fingerprint dimension.  (This process must be repeated for each memory fingerprint dimension in turn.)  These neural predictors will be updated after each recall interval, and will be used during the study period of the next list to select each word to be presented next according to the ongoing neural recordings (Fig.~\ref{fig:adaptiveFR_eeg_order}).

\begin{wrapfigure}[11]{R}{1.7in}
\vspace{-46pt}
  \begin{center}
    \includegraphics[width=1.5in]{figs/adaptive_fr_EEG_flowchart.pdf}
  \end{center}
\vspace{-23pt}
  \caption{\footnotesize \textbf{AdaptiveFR with EEG.}  EEG recordings during study are used to update the memory fingerprint classifier after each batch of recalls.  The classifier is used to adapt the presentation order on a word-by-word basis when the next list is studied.}
  \label{fig:adaptiveFR_eeg_order}
\end{wrapfigure}

\paragraph{Design and procedure.} The experiment and parameters will be analogous to Experiment 2, except that we will use EEG data recorded as participants are studying the words (rather than behavioral data from participants' recalls) to estimate their memory fingerprints.  We plan to record the EEG data using an ActiChamp 160-electrode system (with a photosensor-based electrode mapping system, to allow us to perform source-localization analyses).  Participants will study a total of 16 lists of 16 words each; we will adaptively adjust the presentation orders of each list as described next.

The first four lists will always be sorted randomly.  We will use EEG recordings as participants study these lists to train classifiers to detect the neural signatures of clustering along each memory fingerprint dimension.  Note that, whereas in Experiment 2 the presentation order of the full list is determined before the first word is presented, in Experiment 3 we will be adaptively adjusting the presentation order on a word-by-word basis using classifiers applied to the ongoing EEG recordings.  In other words, we will use the EEG activity recorded during the inter-stimulus interval to predict how the next word will be organized in memory.  Then, using the just-presented word as a reference, we will select a word that is either \textit{consistent}, \textit{inconsistent}, or \textit{random} with respect to that predicted memory fingerprint (Fig.~\ref{fig:adaptiveFR_eeg_order}).  For example, in the \textit{consistent} condition, if the classifier predicts that the next word is likely to be organized by category and font color, then the next to-be-presented word will be chosen to match the category and font color of the just-presented word.  More generally, we will rank all yet-to-be-presented words according to how similar they are along each stimulus dimension (to the just-presented word).  We will then select a matching word by weighting these similarity rankings (using the neurally predicted memory fingerprint to define the weights).  In the \textit{inconsistent} condition, we will next present the word that \textit{least} matches the previous word, using the decoded fingerprint.  In this way, the neurally driven \textit{consistent} and \textit{inconsistent} conditions of Experiment 3 are analogous to the behaviorally driven \textit{stabilize} and \textit{destabilize} conditions of Experiment 2.  We also plan to run yoked control conditions, whereby the presentation orders participants exhibit are driven by \textit{other} participants' neural data.

Using our prior work (Fig.~\ref{fig:clust}) to approximate the relevant parameters, we ran a power analysis (power: 0.8, Type 1 error rate: 0.05) to estimate the number of participants we will need to achieve sufficient power.  The analysis indicates that we will require at least 42 participants in each experimental condition; we plan to run 50 participants using the neurally driven adaptive design described above, plus an additional 50 in the yoked control condition.

\paragraph{Key analyses.}  The overarching goal of Experiment 3 is to test the limits of how the fine-scale temporal dynamics of how people organize incoming information may be leveraged to enhance (or diminish) memory.  The temporal resolution of adaptive learning approaches is limited by the temporal resolution of the measurements on which the adaptive aspects are based.  Whereas recall behaviors may only be assessed periodically (e.g.\ in Experiment 2, during each recall period-- after the participant studies the list), the neural activity driving those behaviors may be measured and analyzed many times per second.  In this way, we seek to use Experiment 3 to answer deep questions about the temporal dynamics of how people organize their memories.  For example, do the same principles about fingerprint stability leading to enhanced memory that we discovered in our pilot studies (Figs.~\ref{fig:random_pilot}, \ref{fig:adaptive_pilot}) also apply to the moment-by-moment fluctuations we will estimate in Experiment 3?  And do people exhibit similar dynamics in those fluctuations within (or across) lists?  Or are those dynamics predictive of recall performance?  In addition to examining the neural estimates of memory fingerprint dynamics, we will also measure the improvement (or detriment) in memory performance across the different experimental conditions (analogous to our proposed analyses in Experiment 2).

Our second set of questions is more general, and is intended to parallel several ideas in Module 2.  Specifically, can we use neural data to uncover (and leverage) the fine-scale temporal dynamics of learning and memory?  In Module 1 (Experiment 3) our focus is on uncovering (and leveraging) the fine-scale temporal dynamics of list learning.  In Module 2 our focus is on uncovering (and leveraging) the fine-scale temporal dynamics of STEM learning.