### 2023-01-02 paper/supplement.tex
finished drafts of all main and supplemental figures

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[font=small,labelfont=bf]{caption}
\usepackage{geometry}
\usepackage{natbib}
\usepackage{pxfonts}
\usepackage{graphicx}
\usepackage{newfloat}
\usepackage{setspace}
\usepackage{hyperref}
\usepackage{placeins}
\usepackage{rotating}

\newcommand{\argmax}{\mathop{\mathrm{argmax}}\limits}
\newcommand{\argmin}{\mathop{\mathrm{argmin}}\limits}

\newcommand{\demo}{1}

\title{\textit{Supplementary materials for}: Carryover effects in free recall reveal how prior experiences influence memories of new experiences}
\author{Jeremy R. Manning\textsuperscript{1, *}, Andrew C. Heusser\textsuperscript{1, 2}, Kirsten Ziman\textsuperscript{1, 3},\\Emily Whitaker\textsuperscript{1}, and Paxton C. Fitzpatrick\textsuperscript{1}\\\textsuperscript{1}Dartmouth College\\\textsuperscript{2}Akili Interactive\\\textsuperscript{3}Princeton University\\\textsuperscript{*}Corresponding author: jeremy.r.manning@dartmouth.edu}
\date{}

\begin{document}

\renewcommand{\figurename}{Supplementary Figure}

%\begin{titlepage}
%  \maketitle
%  \thispagestyle{empty}
%\end{titlepage}

\setcounter{equation}{0}
\setcounter{figure}{0}
\setcounter{table}{0}
\setcounter{page}{1}
\setcounter{section}{0}
\makeatletter
\renewcommand{\theequation}{S\arabic{equation}}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\bibnumfmt}[1]{[S#1]}
\renewcommand{\citenumfont}[1]{S#1}

\maketitle

\begin{figure}[p] \centering
\includegraphics[width=\textwidth]{figures/recall_dynamics_random}

\caption{\textbf{Recall dynamics in feature rich free recall (random conditions).} \textbf{Left panels.} The probabilities of
initiating recall with each word are plotted as a function of presentation
position. \textbf{Middle panels.} The conditional probabilities of recalling
each word are plotted as a function of the relative position (Lag) to the words
recalled just-prior. \textbf{Right panels.} The overall probabilities of
recalling each word are plotted as a function of presentation position.
\textbf{All panels.} Error ribbons denote bootstrap-estimated 95\% confidence
intervals (calculated across participants). Top panels display the recall
dynamics for early (order manipulation) lists in each condition (color). Bottom
panels display the recall dynamics for late (randomly ordered) lists.}

    \label{fig:recall-dynamics-random}
\end{figure}

\begin{figure}[p] \centering
    \includegraphics[width=\textwidth]{figures/recall_dynamics_adaptive}

    \caption{\textbf{Recall dynamics in feature rich free recall (adaptive conditions).} \textbf{Left panels.} The probabilities of
    initiating recall with each word are plotted as a function of presentation
    position. \textbf{Middle panels.} The conditional probabilities of recalling
    each word are plotted as a function of the relative position (Lag) to the words
    recalled just-prior. \textbf{Right panels.} The overall probabilities of
    recalling each word are plotted as a function of presentation position.
    \textbf{All panels.} Error ribbons denote bootstrap-estimated 95\% confidence
    intervals (calculated across participants). Condition is denoted by color.}

        \label{fig:recall-dynamics-adaptive}
    \end{figure}

\begin{figure}[tp] \centering
    \includegraphics[width=0.75\textwidth]{figures/accuracy_by_list}

    \caption{\textbf{Recall accuracy by study list number.}  Each panel displays the average recall accuracy (across participants) as a function of the number of studied lists for the
    random (control) conditions (\textbf{A.}), order manipulation conditions (\textbf{B.}), and adaptive conditions (\textbf{C.}.  The conditions are denoted by color.  Note that the first four
    lists in all of the ``adaptive'' conditions were ordered randomly to compute a baseline fingerprint for each participant prior to initiating the adaptive ordering procedure.  All error
    ribbons denote bootstrap-estimated 95\% confidence intervals, calculated across participants.}
        \label{fig:accuracy-by-list}
\end{figure}

\begin{figure}[tp] \centering
    \includegraphics[width=\textwidth]{figures/clustering_correlations}

    \caption{\textbf{Correlations between feature clustering scores (order manipulation conditions).}  Each column reflects one experimental condition.  The matrices in the top and middle
    rows display across-participant correlations between clustering scores for each feature dimension (top: order manipulation lists; middle: randomly ordered lists).  The matrices
    in the bottom rows display the differences between the top and middle rows.}
        \label{fig:clustering-correlations}
\end{figure}

\begin{figure}[tp] \centering
    \includegraphics[width=\textwidth]{figures/fingerprints_random}

    \caption{\textbf{Memory ``fingerprints'' (random conditions).}
    The across-participant distributions of clustering scores for each feature
    type ($x$-coordinate) are displayed for each experimental condition
    (color), separately for early (top) and late (bottom) lists.}
        \label{fig:fingerprints-random}
\end{figure}

\begin{figure}[tp] \centering
    \includegraphics[width=\textwidth]{figures/fingerprints_adaptive}

    \caption{\textbf{Memory ``fingerprints'' (adaptive conditions).}
    The across-participant distributions of clustering scores for each feature
    type ($x$-coordinate) are displayed for each experimental condition
    (color).}
        \label{fig:fingerprints-adaptive}
\end{figure}

\begin{sidewaysfigure}
    \includegraphics[width=\textwidth]{figures/pnr_matrices}

    \caption{\textbf{Probability of $n$\textsuperscript{th} recall matrices.}  Each sub-panel displays
    the average probability of recalling the given word (Presentation position, $x$-coordinate)
    at the given output position ($y$-coordinate); color denotes the probability.  \textbf{A. Random conditions.}  The top rows
    display data from early (order manipulation) lists, the middle rows display data from late (randomly ordered)
    lists, and the bottom rows display the differences between the matrices in the top and middle rows.  Panel columns denote experimental conditions.
    \textbf{B. Order manipulation conditions.}  The matrices are displayed in the same format as those in Panel A.
    Panel columns are organized by feature type (semantic, lexicographic, or visual).  \textbf{C. Adaptive conditions.}
    The sub-panels are displayed in the same formats as Panels A and B, but here the matrices
    and contrasts (indicated in $y$-axis labels) reflect different list manipulation conditions.}

    \label{fig:pnr}
\end{sidewaysfigure}

\begin{figure}[tp] \centering
    \includegraphics[width=\textwidth]{figures/fingerprint_trajectories_random}

    \caption{\textbf{Memory fingerprint dynamics (random conditions).}
    \textbf{A.} Each column (and color) reflects an experimental condition. In
    the top panels, each marker displays a 2D projection of the
    (across-participant) average memory fingerprint for one list. Early lists
    are denoted by circles and late lists are denoted by stars. All of the
    fingerprints (across all conditions and lists) are projected into a common
    space. The bar plots in the bottom panels display the Euclidean distances
    of the per-list memory fingerprints to the list 0 fingerprint, for each
    condition. Error bars denote bootstrap-estimated 95\% confidence intervals.
    The dotted vertical lines denote the boundaries between early and late
    lists. \textbf{B.} In this panel, the fingerprints for early (circle) and
    late (star) lists are averaged across lists and participants before
    projecting the fingerprints into a (new) 2D space.} \label{fig:fingerprint-trajectories-random}

    \end{figure}

% accuracy by list (list order effects)

% correlations between different types of clustering

%\newpage
%\renewcommand{\refname}{Supplementary references}
%\bibliographystyle{apa}
%\bibliography{CDL-bibliography/cdl}

\end{document}