### 2023-02-03 paper/admin/cover_letter.tex
added draft of cover letter

\title{cover letter}
%
% See http://texblog.org/2013/11/11/latexs-alternative-letter-class-newlfm/
% and http://www.ctan.org/tex-archive/macros/latex/contrib/newlfm
% for more information.
%
\documentclass[11pt,stdletter,orderfromtodate,sigleft]{newlfm}
\usepackage{blindtext, xfrac, animate, hyperref, pxfonts, geometry}

  \setlength{\voffset}{0in}

\newlfmP{dateskipbefore=0pt}
\newlfmP{sigsize=20pt}
\newlfmP{sigskipbefore=10pt}

\newlfmP{Headlinewd=0pt,Footlinewd=0pt}

\newcommand{\journal}{Psychological Review}
\newcommand{\articletype}{Article}
\newcommand{\myTitle}{Feature and order manipulations in a free recall task affect memory for current and future lists}
\newcommand{\corresponding}{Jeremy R. Manning}

\namefrom{\vspace{-0.3in}\corresponding}
\addrfrom{
  Dartmouth College\\
  Department of Psychological and Brain Sciences\\
  HB 6207 Moore Hall\\
  Hanover, NH, 03755}
\addrto{}
\dateset{\today}

\greetto{To the editors of \textit{\journal}:}

\closeline{Sincerely,}
%\usepackage{setspace}
%\linespread{0.85}
% How will your work make others in the field think differently and move the field forward?
% How does your work relate to the current literature on the topic?
% Who do you consider to be the most relevant audience for this work?
% Have you made clear in the letter what the work has and has not achieved?

\begin{document}
\begin{newlfm}

  We have enclosed our manuscript entitled \textit{\myTitle} to be considered
  for publication as an \textit{\articletype}. The experiment reported on in
  our manuscript shows how manipulating the appearances and presentation orders
  of items in a free recall task affects .

  All of the 452 participants in our study encountered the same 16 lists of 16
  words each. The items on each list varied along two semantic dimensions (word
  category, size of referent), two lexicographic dimensions (word length, first
  letter), and two visual dimensions (text color and onscreen presentation
  location). Across different conditions in the experiment, we manipulated
  whether the visual features varied across items (i.e., each word is presented
  in a different font color and at a different location) or whether the visual
  features were held constant across items (i.e., each word is presented in
  black at the center of the screen). We also manipulated the presentation
  orders such that, across different conditions, participants studied some
  lists in a randomized order, and other lists in an order that sorted items
  along one or more feature dimensions. These manipulations changed
  systematically across lists, enabling us to study the consequences of these
  manipulations on the manuplated lists as well as on later lists (subjected to
  potentially different manipulations).

  We observed several particularly exciting findings. First, even though the
  visual features were incidental to the participants' task (i.e., remembering
  as many words as possible), and even though every participant saw exactly the
  same words and lists overall, some of the visual manipulations let
  participants to remember more words. Second, participants showed substantial
  variability in how ``susceptible'' they were to the order manipulations.
  Participants whose behaviors were affected most by the order manipulations
  also showed lingering impacts on later lists that were \textit{not} subjected
  to the same order manipulations. Third, we used a real-time behavioral
  manipulation to change how lists were presented on future lists, according to
  how participants recalled earlier lists. We show that when we organize
  to-be-memorized lists in ways that are compatable with how participants
  ``naturally'' recalled earlier lists, the participants remember more words
  overall. Taken together, our work shows that the \textit{way}
  to-be-remembered content is presented can affect what is remembered, and how.

  We draw several important links between our work and other recent findings
  related to naturalistic memory; context effects on memory performance and
  organization; priming effects on memory; and expectation, event boundaries,
  and situation models. Essentially, we see our manipulations (and findings) as
  helping to bridge key gaps between a large literature describing memory in
  ``traditional'' list-learning tasks, and a growing literature describing
  memory in more ``naturalistic'' settings such as real-world experiences,
  written narratives, movies, and so on. For example, we designed several of
  our experimental manipulations to mimic several aspects of the temporal and
  conceptual structure that are fundamental to naturalistic scenarios. However,
  whereas it can be very difficult to explicitly model or formalize the
  temporal progression of naturalistic experiences, our paradigm is more
  amenable to modeling and incorporating stimulus manipulations.

  We expect that this article will be of interest to readership interested in
  learning and memory, real-time experimental design, priming effects on
  memory, and more.

Thank you for considering our manuscript, and I hope you will find it suitable
for publication in \textit{\journal}.

\end{newlfm}
\end{document}